---
title: "PHYS598 - Data Analysis - March 14, 2019"
author: "Clarisse Vigal"
date: "March 13, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Reading the data in the structure-specific CSV file as an R "dataframe"

```{r}
DVH.CTV7000 <- read.csv("CTV7000.csv", header = TRUE, row.names = NULL)
DVH.CTV5940 <- read.csv("CTV5940.csv", header = TRUE, row.names = NULL)
DVH.Brainstem <- read.csv("Brainstem.csv", header = TRUE, row.names = NULL)
DVH.BrainstemMarg <- read.csv("BrainstemMarg.csv", header = TRUE, row.names = NULL)
DVH.ParotidLT <- read.csv("ParotidLT.csv", header = TRUE, row.names = NULL)
DVH.ParotidRT <- read.csv("ParotidRT.csv", header = TRUE, row.names = NULL)
DVH.Pharynx <- read.csv("Pharynx.csv", header = TRUE, row.names = NULL)
DVH.SpinalCord <- read.csv("SpinalCord.csv", header = TRUE, row.names = NULL)
DVH.SpinalCordMarg <- read.csv("SpinalCordMarg.csv", header = TRUE, row.names = NULL)
DVH.SubmandLT <- read.csv("SubmandLT.csv", header = TRUE, row.names = NULL)
DVH.SubmandRT <- read.csv("SubmandRT.csv", header = TRUE, row.names = NULL)
```

Most important structures with specific dose parameters

```{r}
CTV7000_D95 = DVH.CTV7000[,c(1,3,98)]
CTV7000_D99 = DVH.CTV7000[,c(1,3,102)]
CTV7000_D20 = DVH.CTV7000[,c(1,3,23)]
CTV5940_D99 = DVH.CTV5940[,c(1,3,102)]
CTV5940_D95 = DVH.CTV5940[,c(1,3,98)]
CTV5940_D20 = DVH.CTV5940[,c(1,3,23)]
Brainstem_D1 = DVH.Brainstem[,c(1,3,4)]
ParotidLT_D50 = DVH.ParotidLT[,c(1,3,53)]
ParotidRT_D50 = DVH.ParotidRT[,c(1,3,53)]
SpinalCord_D1 = DVH.SpinalCord[,c(1,3,4)]
```

Obtaining the row numbers with new patient

```{r}
#Deleting NA entries
CTV7000_D99 <- CTV7000_D99[complete.cases(CTV7000_D99),]
CTV5940_D99 <- CTV5940_D99[complete.cases(CTV5940_D99),]
Brainstem_D1 <- Brainstem_D1[complete.cases(Brainstem_D1),]
ParotidLT_D50 <- ParotidLT_D50[complete.cases(ParotidLT_D50),]
ParotidRT_D50 <- ParotidRT_D50[complete.cases(ParotidRT_D50),]
SpinalCord_D1 <- SpinalCord_D1[complete.cases(SpinalCord_D1),]

CTV7000_ptrows<- c(1);
for (i in 1:(length(CTV7000_D99[,1])-1))
{
  if  (!(CTV7000_D99[i,1] == CTV7000_D99[i+1,1]))
    {CTV7000_ptrows <- c(CTV7000_ptrows,i+1)}
}

CTV5940_ptrows<- c(1);

for (i in 1:(length(CTV5940_D99[,1])-1))
{
  if  (!(CTV5940_D99[i,1] == CTV5940_D99[i+1,1]))
    {
    CTV5940_ptrows <- c(CTV5940_ptrows,i+1)}
}

Brainstem_ptrows<- c(1);

for (i in 1:(length(Brainstem_D1[,1])-1))
{
  if  (!(Brainstem_D1[i,1] == Brainstem_D1[i+1,1]))
    {
    Brainstem_ptrows <- c(Brainstem_ptrows,i+1)}
}

ParotidLT_ptrows<- c(1);

for (i in 1:(length(ParotidLT_D50[,1])-1))
{
  if  (!(ParotidLT_D50[i,1] == ParotidLT_D50[i+1,1]))
    {
    ParotidLT_ptrows <- c(ParotidLT_ptrows,i+1)}
}

ParotidRT_ptrows<- c(1);

for (i in 1:(length(ParotidRT_D50[,1])-1))
{
  if  (!(ParotidRT_D50[i,1] == ParotidRT_D50[i+1,1]))
    {
    ParotidRT_ptrows <- c(ParotidRT_ptrows,i+1)}
}

SpinalCord_ptrows<- c(1);

for (i in 1:(length(SpinalCord_D1[,1])-1))
{
  if  (!(SpinalCord_D1[i,1] == SpinalCord_D1[i+1,1]))
    {
    SpinalCord_ptrows <- c(SpinalCord_ptrows,i+1)}
}

# Max function for tables with NA
my.max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA)
my.min <- function(x) ifelse( !all(is.na(x)), min(x, na.rm=T), NA)

```


```{r}
library("RColorBrewer")
#Initializing a color vector to help us color code our graphs for each patient
color = brewer.pal(n = 10,name = "Spectral")
color = colorRampPalette(color)(60)
```

Plotting CTV7000 D95% in time

```{r}
plot(NULL,xlab="Fraction Number (/33)", ylab="CTV7000 D95%",xlim = c(0,33),ylim = c(my.min(CTV7000_D95$D95.Gy) - 2,my.max(CTV7000_D95$D95.Gy) + 2))

#Iterating through the patients and plotting trends in D95% with fraction number
for (i in 1:(length(CTV7000_ptrows)-1) )
{
  rm(vec)
  ind1 = CTV7000_ptrows[i]
  ind2 = CTV7000_ptrows[i+1]-1
  vec <- data.frame(Fraction.Number = CTV7000_D95[c(ind1:ind2),2], D95.Gy = CTV7000_D95[c(ind1:ind2),3])
  points(vec$Fraction.Number,vec$D95.Gy,pch = 19, col = color[i])
  lines(vec$Fraction.Number,vec$D95.Gy,pch = 19, col = color[i])
}
```

* more or less consistent

Brainstem D1%

```{r}
plot(NULL,xlab="Fraction Number (/32)", ylab="Brainstem D1%",xlim = c(0,32),ylim = c(my.min(Brainstem_D1$D1.Gy) - 2,my.max(Brainstem_D1$D1.Gy) + 2))

#Iterating through the patients and plotting trends with fraction number
for (i in 1:(length(Brainstem_ptrows)-1) )
{
  rm(vec)
  ind1 = Brainstem_ptrows[i]
  ind2 = Brainstem_ptrows[i+1]-1
  vec <- data.frame(Fraction.Number = Brainstem_D1[c(ind1:ind2),2], D1.Gy = Brainstem_D1[c(ind1:ind2),3])
  points(vec$Fraction.Number,vec$D1.Gy,pch = 19, col = color[i])
   lines(vec$Fraction.Number,vec$D1.Gy,pch = 19, col = color[i])
}
```

also consistent

Parotid LT D50%

```{r}
plot(NULL,xlab="Fraction Number (/33)", ylab="Dose Distribution (Gy)", main = "Parotid LT D50%",xlim = c(0,33),ylim = c(my.min(ParotidLT_D50$D50.Gy) - 2,my.max(ParotidLT_D50$D50.Gy) + 2))

#Iterating through the patients and plotting trends with fraction number
for (i in 1:(length(ParotidLT_ptrows)-1) )
{
  rm(vec)
  ind1 = ParotidLT_ptrows[i]
  ind2 = ParotidLT_ptrows[i+1]-1
  vec <- data.frame(Fraction.Number = ParotidLT_D50[c(ind1:ind2),2], D50.Gy = ParotidLT_D50[c(ind1:ind2),3])
  points(vec$Fraction.Number,vec$D50.Gy,pch = 19, col = color[i])
   lines(vec$Fraction.Number,vec$D50.Gy,pch = 19, col = color[i])
}
```

Parotid RT D50%

```{r}
library("RColorBrewer")
#Initializing a color vector to help us color code our graphs for each patient
color = brewer.pal(n = 10,name = "Spectral")
color = colorRampPalette(color)(60)

#R needs a generic "plot" call before "points" or "lines" can be superimposed

plot(NULL,xlab="Fraction Number (/33)", ylab="Dose Distribution (Gy)", main = "Parotid RT D50%",xlim = c(0,33),ylim = c(my.min(ParotidRT_D50$D50.Gy) - 2,my.max(ParotidRT_D50$D50.Gy) + 2))

#Iterating through the patients and plotting trends with fraction number
for (i in 1:(length(ParotidRT_ptrows)-1) )
{
  rm(vec)
  ind1 = ParotidRT_ptrows[i]
  ind2 = ParotidRT_ptrows[i+1]-1
  vec <- data.frame(Fraction.Number = ParotidRT_D50[c(ind1:ind2),2], D50.Gy = ParotidRT_D50[c(ind1:ind2),3])
  points(vec$Fraction.Number,vec$D50.Gy,pch = 19, col = color[i])
   lines(vec$Fraction.Number,vec$D50.Gy,pch = 19, col = color[i])
}
```

* both parotids are changing quite a bit in time.


### Plotting Full DVH plots in 2D
- to visualize the spread, and what changes look like beyond just conventional Dxx%

```{r}
library("RColorBrewer")
#Initializing a color vector to help us color code our graphs for each patient
color = brewer.pal(n = 5,name = "Dark2")
color = colorRampPalette(color)(5)

DVHplot.ParotidRT <- function(x,y)
{
  plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",xlim = c(0,100), ylim = c(my.min(DVH.ParotidRT[,c(4:103)]) - 2,my.max(DVH.ParotidRT[,c(4:103)]) + 2), main="2D Graph of Patient-Specific DVHS for Parotid RT")
  for (i in x:y )
{
  for (j in 0:(ParotidRT_ptrows[i+1]-ParotidRT_ptrows[i] - 1) )
  {
    ind1 = ParotidRT_ptrows[i] + j
    lines(seq(1,100,1),DVH.ParotidRT[ind1,c(4:103)],pch = 19, col = color[i-x+1])
  }
}
}

DVHplot.ParotidLT <- function(x,y)
{
  plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",xlim = c(0,100), ylim = c(my.min(DVH.ParotidLT[,c(4:103)]) - 2,my.max(DVH.ParotidLT[,c(4:103)]) + 2), main="2D Graph of Patient-Specific DVHS for Parotid LT")
  for (i in x:y )
{
  for (j in 0:(ParotidLT_ptrows[i+1]-ParotidLT_ptrows[i] - 1) )
  {
    ind1 = ParotidLT_ptrows[i] + j
    lines(seq(1,100,1),DVH.ParotidLT[ind1,c(4:103)],pch = 19, col = color[i-x+1])
  }
}
}
DVHplot.CTV7000 <- function(x,y)
{
  plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",xlim = c(0,100), ylim = c(my.min(DVH.CTV7000[,c(4:103)]) - 2,my.max(DVH.CTV7000[,c(4:103)]) + 2), main="2D Graph of Patient-Specific DVHS for CTV7000")
  for (i in x:y )
{
  for (j in 0:(CTV7000_ptrows[i+1]-CTV7000_ptrows[i] - 1) )
  {
    ind1 = CTV7000_ptrows[i] + j
    lines(seq(1,100,1),DVH.CTV7000[ind1,c(4:103)],pch = 19, col = color[i-x+1])
  }
}
}
```

Looking at CTV7000 again:

```{r}
DVHplot.CTV7000(1,5)
DVHplot.CTV7000(6,10)
DVHplot.CTV7000(11,15)
DVHplot.CTV7000(16,20)
DVHplot.CTV7000(21,25)
DVHplot.CTV7000(26,30)
DVHplot.CTV7000(31,35)
DVHplot.CTV7000(36,40)
DVHplot.CTV7000(41,45)
DVHplot.CTV7000(46,50)
DVHplot.CTV7000(51,55)
DVHplot.CTV7000(56,59)
```

Parotid RT

```{r}
DVHplot.ParotidRT(1,5)
DVHplot.ParotidRT(6,10)
DVHplot.ParotidRT(11,15)
DVHplot.ParotidRT(16,20)
DVHplot.ParotidRT(21,25)
DVHplot.ParotidRT(26,30)
DVHplot.ParotidRT(31,35)
DVHplot.ParotidRT(36,40)
DVHplot.ParotidRT(41,45)
DVHplot.ParotidRT(46,50)
DVHplot.ParotidRT(51,55)
DVHplot.ParotidRT(56,58)
```

Parotid LT

```{r}
DVHplot.ParotidLT(1,5)
DVHplot.ParotidLT(6,10)
DVHplot.ParotidLT(11,15)
DVHplot.ParotidLT(16,20)
DVHplot.ParotidLT(21,25)
DVHplot.ParotidLT(26,30)
DVHplot.ParotidLT(31,35)
DVHplot.ParotidLT(36,40)
DVHplot.ParotidLT(41,45)
DVHplot.ParotidLT(46,50)
DVHplot.ParotidLT(51,55)
DVHplot.ParotidLT(56,58)
```

### Plotting Lag 1 Difference
- to visualize whether there is a systematic trend

```{r}
lagParotidRT <- function(x,y) {
  
maxlist <- c(0)
minlist <- c(0)
for (i in x:y) #first 5 patients
{
  for (j in 0:(ParotidRT_ptrows[i+1]-ParotidRT_ptrows[i] - 2) ) #time index
  {
    ind1 = ParotidRT_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.ParotidRT[ind1,c(4:103)]
    t2 = DVH.ParotidRT[ind2,c(4:103)]
    lagdiff = t1-t2
    maxlist <- c(maxlist,max(lagdiff))
    minlist <- c(minlist,min(lagdiff))
  }

}

plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",main = "Lag 1 Differencing for Parotid RT",xlim = c(0,100), ylim = c(min(minlist),max(maxlist)))
for (i in x:y) 
{
  for (j in 0:(ParotidRT_ptrows[i+1]-ParotidRT_ptrows[i] - 2) ) #time index
  {
    ind1 = ParotidRT_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.ParotidRT[ind1,c(4:103)]
    t2 = DVH.ParotidRT[ind2,c(4:103)]
    lagdiff = t1-t2
    lines(seq(1,100,1),lagdiff, col = color[i-x+1])
    
      }
}

}

lagParotidLT <- function(x,y) {
  
maxlist <- c(0)
minlist <- c(0)
for (i in x:y) #first 5 patients
{
  for (j in 0:(ParotidLT_ptrows[i+1]-ParotidLT_ptrows[i] - 2) ) #time index
  {
    ind1 = ParotidLT_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.ParotidLT[ind1,c(4:103)]
    t2 = DVH.ParotidLT[ind2,c(4:103)]
    lagdiff = t1-t2
    maxlist <- c(maxlist,max(lagdiff))
    minlist <- c(minlist,min(lagdiff))
  }

}
#print(max(maxlist))
plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",main = "Lag 1 Differencing for Parotid LT",xlim = c(0,100), ylim = c(min(minlist),max(maxlist)))#,asp = 2)
for (i in x:y) #first 5 patients
{
  for (j in 0:(ParotidLT_ptrows[i+1]-ParotidLT_ptrows[i] - 2) ) #time index
  {
    ind1 = ParotidLT_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.ParotidLT[ind1,c(4:103)]
    t2 = DVH.ParotidLT[ind2,c(4:103)]
    lagdiff = t1-t2
    lines(seq(1,100,1),lagdiff, col = color[i-x+1])
    
      }
}


}

lagCTV7000 <- function(x,y) {
  
maxlist <- c(0)
minlist <- c(0)
for (i in x:y) #first 5 patients
{
  for (j in 0:(CTV7000_ptrows[i+1]-CTV7000_ptrows[i] - 2) ) #time index
  {
    ind1 = CTV7000_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.CTV7000[ind1,c(4:103)]
    t2 = DVH.CTV7000[ind2,c(4:103)]
    lagdiff = t1-t2
    maxlist <- c(maxlist,max(lagdiff))
    minlist <- c(minlist,min(lagdiff))
  }

}
#print(max(maxlist))
plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",main = "Lag 1 Differencing for CTV7000",xlim = c(0,100), ylim = c(min(minlist),max(maxlist)))#,asp = 2)
for (i in x:y) #first 5 patients
{
  for (j in 0:(CTV7000_ptrows[i+1]-CTV7000_ptrows[i] - 2) ) #time index
  {
    ind1 = CTV7000_ptrows[i] + j
    ind2 = ind1 + 1
    t1 = DVH.CTV7000[ind1,c(4:103)]
    t2 = DVH.CTV7000[ind2,c(4:103)]
    lagdiff = t1-t2
    lines(seq(1,100,1),lagdiff, col = color[i-x+1])
    
      }
}


}
```

Lag 1 for CTV7000


```{r}
lagCTV7000(1,5)
lagCTV7000(6,10)
lagCTV7000(11,15)
lagCTV7000(16,20)
lagCTV7000(21,25)
lagCTV7000(26,30)
lagCTV7000(31,35)
lagCTV7000(36,40)
lagCTV7000(41,45)
lagCTV7000(46,50)
lagCTV7000(51,55)
lagCTV7000(56,59)
```

Lag 1 for Parotid RT

```{r}
lagParotidRT(1,5)
lagParotidRT(6,10)
lagParotidRT(11,15)
lagParotidRT(16,20)
lagParotidRT(21,25)
lagParotidRT(26,30)
lagParotidRT(31,35)
lagParotidRT(36,40)
lagParotidRT(41,45)
lagParotidRT(46,50)
lagParotidRT(51,55)
lagParotidRT(56,58)
```

```{r}
lagParotidLT(1,5)
lagParotidLT(6,10)
lagParotidLT(11,15)
lagParotidLT(16,20)
lagParotidLT(21,25)
lagParotidLT(26,30)
lagParotidLT(31,35)
lagParotidLT(36,40)
lagParotidLT(41,45)
lagParotidLT(46,50)
lagParotidLT(51,55)
lagParotidLT(56,58)
```

### Vertical Translation
- to visualize random effects - random noise?


```{r}
library("RColorBrewer")
#Initializing a color vector to help us color code our graphs for each patient
color = brewer.pal(n = 10,name = "Spectral")
color = colorRampPalette(color)(60)

verticaltranslation.ParotidRT <- function(x,y)
{
plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)", main = "Vertical Translation for Parotid RT" ,xlim = c(0,100), ylim = c(0,30))#,asp = 2)

for (i in x:y)
{
  vtrans <- c()
  for (j in 4:103)
  {
    ind1 = ParotidRT_ptrows[i]
    ind2 = ParotidRT_ptrows[i+1] - 1
    mindose = min(DVH.ParotidRT[ind1:ind2,j])
    maxdose = max(DVH.ParotidRT[ind1:ind2,j])
    trans = maxdose-mindose
    vtrans <- c(vtrans,trans)
    
  }
  lines(seq(1,100,1),vtrans, col = color[i-x+1])
}
}

verticaltranslation.ParotidLT <- function(x,y)
{
plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)", main = "Vertical Translation for Parotid LT" ,xlim = c(0,100), ylim = c(0,30))#,asp = 2)

for (i in x:y)
{
  vtrans <- c()
  for (j in 4:103)
  {

    ind1 = ParotidLT_ptrows[i]
    ind2 = ParotidLT_ptrows[i+1] - 1
    mindose = min(DVH.ParotidLT[ind1:ind2,j])
    maxdose = max(DVH.ParotidLT[ind1:ind2,j])
    trans = maxdose-mindose
    vtrans <- c(vtrans,trans)
    
  }
  lines(seq(1,100,1),vtrans, col = color[i-x+1])
}
}

verticaltranslation.CTV7000 <- function(x,y)
{
plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)", main = "Vertical Translation for CTV7000" ,xlim = c(0,100), ylim = c(0,45))#,asp = 2)

for (i in x:y)
{
  vtrans <- c()
  for (j in 4:103)
  {
    #mindose = min(DVH.CTV7000[CTV7000_ptrows[i]:CTV7000_ptrows[i+1]],j)
    #dvol = DVH.CTV7000[CTV7000_ptrows[i]:CTV7000_ptrows[i+1],j]
    ind1 = CTV7000_ptrows[i]
    ind2 = CTV7000_ptrows[i+1] - 1
    #print(ind2)
    mindose = min(DVH.CTV7000[ind1:ind2,j])
    maxdose = max(DVH.CTV7000[ind1:ind2,j])
    trans = maxdose-mindose
    vtrans <- c(vtrans,trans)
    
  }
  lines(seq(1,100,1),vtrans, col = color[i-x+1])
}
}

```

```{r}
verticaltranslation.CTV7000(1,59)
verticaltranslation.ParotidLT(1,58)
verticaltranslation.ParotidRT(1,58)
```

### Taking the slope and standard error of each dose parameter with resepct with time

Function that looks at each dose parameter separately and fits a line corresponding to the change in that parameter with respect to time.
Return Values:
1. Maximum slope of a line of best fit when considering all dose parameters - adaptive radiaton therapy related, greatest systematic trend
2. Index of the maximum slope
3. Maximum standard error of a line of best fit when considering all dose considering all dose parameters - daily set up realted, greatest random noise
4. Index of the maximum standard error

### Parotid RT

```{r}
featuresParotidRT <- function(i)
{

vecSlope = rep(NA,100)
vecSlopeSE = rep(NA,100)

numEl = ParotidRT_ptrows[i+1] - ParotidRT_ptrows[i]
vecInd = c(ParotidRT_ptrows[i]:(ParotidRT_ptrows[i+1] - 1))

for (j in 1:100)
  {
  vecDoseParam = DVH.ParotidRT[vecInd,3+j]
  vecFraction = DVH.ParotidRT[vecInd,3]
  # Weighting the first observation (CTsim) more heavily to ensure line of best fit passess through it
  vecWeights = rep(1,numEl)
  vecWeights[1] <- 100
  lm.fit = lm(vecDoseParam ~ vecFraction, weights = vecWeights)
  #lm.fit = lm(vecDoseParam ~ vecFraction)
  
  #plot(NULL,xlab="Structure Volume (%)", ylab="Dose Distribution (Gy)",xlim = c(0,33), ylim = c(my.min(DVH.ParotidRT[vecInd,c(3+j)]) - 2,my.max(DVH.ParotidRT[vecInd,c(3+j)]) + 2), main="2D Graph of Temporal Dose Parameter Change for Parotid RT")
  #points(vecFraction, vecDoseParam, col = "blue")
  #abline(lm.fit,lwd = 3, col = "red")
  
  #vecIntercept[j] = coef(summary(lm.fit))[1,1]
  vecSlope[j] = coef(summary(lm.fit))[2,1]
  vecSlopeSE[j] = coef(summary(lm.fit))[2,2]
}

maxSlope = max(vecSlope)
maxSlopeInd = which.max(vecSlope)
maxSlopeSE = max(vecSlopeSE)
maxSlopeSEInd = which.max(vecSlopeSE)

plot(NULL,xlab="Fraction Number", ylab="Dose Parameter Value (Gy)",xlim = c(0,33), ylim = c(my.min(DVH.ParotidRT[vecInd,maxSlopeInd+3]) - 2,my.max(DVH.ParotidRT[vecInd,maxSlopeInd+3]) + 2), main="2D Graph of Temporal Dose Parameter Change for Parotid RT (at max slope)")
points(vecFraction, DVH.ParotidRT[vecInd,maxSlopeInd+3], col = "blue")
lmMax.fit = lm(DVH.ParotidRT[vecInd,maxSlopeInd+3] ~ vecFraction, weights = vecWeights)
abline(lmMax.fit,lwd = 3, col = "red")

returnVec = c(maxSlope, maxSlopeInd, maxSlopeSE, maxSlopeSEInd)

}
```

```{r}
library("openxlsx")
numRows = length(ParotidRT_ptrows) - 1
ptMat_toExport = matrix(nrow = numRows, ncol = 4)
for (i in 1:numRows){
  ptMat_toExport[i,] = featuresParotidRT(i)
}
write.xlsx(ptMat_toExport,file = "exportTest_ParotidRT.xlsx")
```

Using the data extracted and ML methods:

```{r}
library(tree)

rm(list = ls())

dataTypes = c("integer","integer","factor","integer","factor",
              "factor","factor","factor","factor","factor",
              "factor","factor","numeric","numeric","factor",
              "factor","numeric","numeric","numeric","numeric",
              "numeric","numeric","numeric","numeric","numeric",
              "numeric","numeric","numeric","integer","numeric",
              "integer")

Data_ParotidRT.Full <- read.csv("AnonymizedData_forML.csv", header = TRUE, 
                    colClasses = dataTypes, row.names = NULL) 

# Min and Max function for tables with NA
my.max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA)
my.min <- function(x) ifelse( !all(is.na(x)), min(x, na.rm=T), NA)
```

```{r}
#Domain-based Feature Selection
Data_ParotidRT.Select = Data_ParotidRT.Full[c(1:34,36:60),c(5:8,13:14,23,28:31)] 
```

### Clustering 

```{r}
plot(NULL,xlab="Index", ylab="Slope",xlim = c(0,100), ylim = c(my.min(Data_ParotidRT.Select[,8]), my.max(Data_ParotidRT.Select[,8])), main="Index and Value of Maximum Slope for Parotid RT")
points(Data_ParotidRT.Select[,9], Data_ParotidRT.Select[,8], col = "blue")

plot(NULL,xlab="Index", ylab="Slope",xlim = c(0,100), ylim = c(my.min(Data_ParotidRT.Select[,10]), my.max(Data_ParotidRT.Select[,10])), main="Index and Value of Maximum SlopeSE for Parotid RT")
points(Data_ParotidRT.Select[,11], Data_ParotidRT.Select[,10], col = "blue")
```

```{r}

##All Data
Predictors = Data_ParotidRT.Select[,c(1:7)]
Response = Data_ParotidRT.Select[,c(9)]

set.seed(2)
train = sample(1:nrow(Predictors),35)
Predictors.train = Predictors[train,]
Predictors.test = Predictors[-train,]
Response.train = Response[train]
Response.test = Response[-train]
```

### Simple Classification Tree

```{r}
result.tree = tree(Response.train~.,Predictors.train)
summary(result.tree)
plot(result.tree)
text(result.tree,pretty=0)
```






